---
title: "Project Results"
author: "Devin Luu, DTL527"
output: html_document
---
The amount of $NO_{2}$ found in various areas have strong correlations to several predictors. We have found that the strongest predictors of $NO_{2}$
pollution comes from a combination of the area's population and ozone ($O_{3}$) pollution.

```{r include=FALSE}
library(tidyverse)
library(modelr)
library(data.world)
library(scatterplot3d)

options(na.action = na.warn)

project <- "https://data.world/mattstras/f-17-eda-project-1"
df <- data.world::query(
  data.world::qry_sql("SELECT * FROM airQualityData"),
  dataset = project
)
attach(df)
```

```{r}

lmModel1 <- lm(no2_1_hr_ppb ~ log_pop_2010 + o3_8_hr_ppm)
summary(lmModel1)

lmPlot <- scatterplot3d(log_pop_2010,o3_8_hr_ppm,no2_1_hr_ppb, xlab = "log(population)", ylab = "O3, ppm", zlab = "no2, hourly, ppb", pch = 16, color = "blue", angle = 330)
lmPlot$plane3d(lmModel1,col = "red")
```

We can estimate that an increase of $10^{population}$ will contribute to $NO_{2}$ hourly measurements by 12.7 ppb and an increase of ozone by 1 ppm will contribute 628 ppb.

The residuals from this model are evenly dispersed and have no particular slant or shape indicative of underlying effects.

```{r}
residDF <- df %>% select(log_pop_2010, o3_8_hr_ppm, no2_1_hr_ppb)

residDF <- residDF %>% add_residuals(lmModel1, "resid")

resPlot <- scatterplot3d(x = residDF$log_pop_2010, y = residDF$o3_8_hr_ppm, z = residDF$resid, xlab = "log(population)", ylab = "O3, ppm", zlab = "residuals", pch =16, color = "blue", angle = 200)
resPlot$plane3d(c(0,0,0), col = "red")
```

Using the same model to predict daily $NO_{2}$ measurements produced similar results with $R^{2} = 0.682$.

```{r}
lmModel12 <- lm(no2_am_ppb ~ log_pop_2010 + o3_8_hr_ppm)
summary(lmModel12)

lmPlot <- scatterplot3d(log_pop_2010,o3_8_hr_ppm,no2_am_ppb, xlab = "log(population)", ylab = "O3, ppm", zlab = "no2, daily, ppb", pch = 16, color = "blue", angle = 330)
lmPlot$plane3d(lmModel12,col = "red")

residDF <- df %>% select(log_pop_2010, o3_8_hr_ppm, no2_am_ppb)

residDF <- residDF %>% add_residuals(lmModel12, "resid")

resPlot <- scatterplot3d(x = residDF$log_pop_2010, y = residDF$o3_8_hr_ppm, z = residDF$resid, xlab = "log(population)", ylab = "O3, ppm", zlab = "residuals", pch =16, color = "blue", angle = 200)
resPlot$plane3d(c(0,0,0), col = "red")
```

While the both models have identical predictor terms, the $R^{2}$ value for the daily model is slightly less than that of the hourly. This might have something to do with the nonlinear relationship between the two types of measurements.

```{r}
lmModel17 <- lm(no2_am_ppb ~ no2_1_hr_ppb + I(no2_1_hr_ppb^2))
summary(lmModel17)

grid4 <- df %>% data_grid(no2_1_hr_ppb = seq_range(no2_1_hr_ppb,20)) %>% add_predictions(lmModel17, "no2_am_ppb")


ggplot(df, mapping=aes(no2_1_hr_ppb, no2_am_ppb)) + geom_point() +  geom_line(data = grid4, col="red", lwd = 1)

res4 <- df %>% add_residuals(lmModel17,"resids")
ggplot(res4, mapping = aes(no2_1_hr_ppb, resids)) + geom_point() + geom_hline(yintercept = 0, col = "red")

```

The residual plot for this model shows heteroscedasticity so instead we will look at the log transform of the daily measurements.

```{r}
transDf <- df %>% mutate(log_no2_am_ppb = log(no2_am_ppb))
ggplot(transDf, mapping = aes(no2_1_hr_ppb, log_no2_am_ppb)) + geom_point()

lmModel17_2 <- lm(log_no2_am_ppb ~ no2_1_hr_ppb + I(no2_1_hr_ppb^2), data = transDf)
summary(lmModel17_2)

grid5 <- transDf %>% data_grid(no2_1_hr_ppb = seq_range(no2_1_hr_ppb,20)) %>% add_predictions(lmModel17_2, "log_no2_am_ppb")

ggplot(transDf, mapping = aes(no2_1_hr_ppb, log_no2_am_ppb)) + geom_point() + geom_line(data = grid5, col = "red")

res5 <- transDf %>% add_residuals(lmModel17_2,"resids")
ggplot(res5, mapping = aes(no2_1_hr_ppb, resids)) + geom_point() + geom_hline(yintercept = 0, col = "red")


```
